\section{Method}
We adopt the method from RALA for solving the low rank problem, and replace the weighting component with an RL weighting framework which is a continuous, stochastic Actor-Critic module. We aim to add a budget constraint, and layer wise information into the weighting to improve the linear attention speed and performance. 
\subsection{Architecture Overview} 
An input image is tokenized into $N$ patch embeddings and partitioned into $T$ sequential chunks of size $C$. A linear-attention ViT backbone processes each chunk, projecting it into local queries $Q_t$, keys $K_t$, and values $V_t$ and maintaining a recurrent memory $(S_t, Z_t)$ in place of a full attention matrix. At each step $t$, an Actor-Critic module observes the causal state local features, compressed memory, and remaining budget—and emits per-token weights $w_t \in (0,1)^C$. These weights gate the memory update $S_t = S_{t-1} + (w_t \odot \phi(K_t))^\top V_t$. After all $T$ chunks, the final state $(S_T, Z_T)$ is consumed by a task head (classifier or detection decoder), which produces the terminal reward.
 
Training follows a two-phase curriculum. \textbf{Phase~1} bootstraps the routing module from Fisher Information saliency targets, avoiding the policy collapse that arises when RL is applied cold to a high-dimensional visual state. \textbf{Phase~2} fine-tunes the full system end-to-end on the downstream task via PPO with Generalized Advantage Estimation. During inference,  stochastic sampling is disabled, the Actor's deterministic outputs are hard-thresholded at $\tau$, physically pruning low-weight tokens from the GPU buffer and yielding exact FLOP reductions.

\begin{figure*}
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[
    font=\small,
    >={Stealth[scale=1.1]},
    node distance=1.4cm and 1.6cm,
    block/.style={
      draw, rounded corners=4pt,
      fill=blue!6, draw=blue!30,
      minimum width=2.8cm, minimum height=0.85cm,
      align=center, inner sep=4pt
    },
    phase/.style={
      draw, rounded corners=4pt,
      fill=gray!7, draw=gray!40,
      minimum width=2.8cm, minimum height=0.85cm,
      align=center, inner sep=4pt
    },
    actor/.style={
      draw, rounded corners=4pt,
      fill=red!8, draw=red!35,
      minimum width=3.8cm, minimum height=1.1cm,
      align=center, inner sep=5pt
    },
    memory/.style={
      draw, cylinder, shape border rotate=90, aspect=0.22,
      fill=green!8, draw=green!40,
      minimum height=1.4cm, minimum width=2.6cm,
      align=center
    },
    gate/.style={
      draw, rounded corners=4pt,
      fill=orange!10, draw=orange!40,
      minimum width=2.8cm, minimum height=0.85cm,
      align=center, inner sep=4pt
    },
    head/.style={
      draw, rounded corners=4pt,
      fill=purple!8, draw=purple!35,
      minimum width=3.0cm, minimum height=0.9cm,
      align=center, inner sep=4pt
    },
    reward/.style={
      draw, rounded corners=4pt,
      fill=yellow!12, draw=yellow!50!black,
      minimum width=2.8cm, minimum height=0.85cm,
      align=center, inner sep=4pt
    },
    annot/.style={font=\scriptsize, text=gray!70},
    larrow/.style={->, thick},
    darrow/.style={->, dashed, gray!60},
    rarrow/.style={->, thick, dashed, green!50!black}
  ]

  % ---- ROW 0: input pipeline ----
  \node[block]  (img)    at (0,   0) {Input Image};
  \node[block]  (patch)  at (4.5, 0) {Patch Embed\\$N$ tokens};
  \node[block]  (chunk)  at (9,   0) {Chunk $t$\\$K_t,Q_t,V_t\in\mathbb{R}^{C\times D}$};
  \node[phase]  (gctx)   at (14,  0) {Running context\\$G_t=Z_{t-1}/(t{-}1)$};

  \draw[larrow] (img)   -- (patch);
  \draw[larrow] (patch) -- (chunk);
  \draw[larrow] (chunk) -- (gctx);

  % ---- ROW 1: memory + state assembly ----
  \node[memory] (mem_prev) at (0,   -3.5) {Memory\\$S_{t-1},Z_{t-1}$};
  \node[block]  (state)    at (9,   -3.5) {Causal State $s_t$\\
    $[K_t \parallel \mathrm{Proj}(S_{t-1})$\\
    $\parallel\,\mathrm{Proj}(Z_{t-1})\parallel B_t]$};
  \node[phase]  (budget)   at (14, -3.5) {Budget\\$B_t\in[0,1]$};

  \draw[larrow] (chunk)    -- (state);
  \draw[larrow] (mem_prev) -- (state);
  \draw[larrow] (budget)   -- (state);
  \draw[darrow] (gctx.south) -- ++(0,-0.6) -| (state.north);

  % ---- ROW 2: Actor-Critic ----
  \node[actor] (actor)  at (9,   -7.0)
    {PPO Actor-Critic $\pi_\theta$\\
     $w_t\sim\mathrm{SquashedGaussian}(\mu_\theta,\sigma_\theta)$};
  \node[phase] (fisher) at (3.0, -7.0)
    {Phase 1: FI init\\
     $\mathcal{L}_\text{A}{=}\mathbb{E}[(w_\mu{-}\tilde{\mathcal{I}})^2]$};

  \draw[larrow] (state)  -- (actor);
  \draw[darrow, <->] (fisher) --
    node[annot, above]{warm-start} (actor);

  \node[annot, below=0.18cm of actor] (wannot) {$w_t\in(0,1)^C$};

  % ---- ROW 3: gating split ----
  \node[gate] (soft) at (6.5, -10.0)
    {Soft Scale\\$\hat{K}_t = w_t\odot\phi(K_t)$};
  \node[gate] (hard) at (11.5,-10.0)
    {Hard Prune (infer.)\\$\mathbb{I}[w_t>\tau]$};

  \draw[larrow] (actor) -| (soft);
  \draw[larrow] (actor) -| (hard);

  % ---- ROW 4: dilution + active tokens ----
  \node[gate]  (dilution) at (6.5, -13.0)
    {State Dilution\\
     $+\,\gamma(1{-}\bar{w}_t)\cdot S_{t-1}/(t{-}1)$};
  \node[block] (active)   at (11.5,-13.0)
    {Active tokens\\$C_\mathrm{active}\leq C$};

  \draw[larrow] (soft) -- (dilution);
  \draw[larrow] (hard) -- (active);

  % ---- ROW 5: updated memory ----
  \node[memory] (mem_new) at (9, -16.0)
    {Updated Memory\\$S_t,\,Z_t$};

  \draw[larrow] (dilution) |- (mem_new);
  \draw[larrow] (active)   |- (mem_new);

  \draw[rarrow] (mem_prev.south) -- ++(0,-1.0)
    -| node[annot, pos=0.18, below]{recurrent} (mem_new.west);

  % budget update node
  \node[phase] (budgetup) at (14.5,-16.0)
    {$B_{t+1}=B_t-\tfrac{1}{N}\sum w_{i,t}$};
  \draw[larrow] (mem_new) -- (budgetup);

  % recurrence arrow back to chunk
  \draw[rarrow, thick]
    (budgetup.north) -- ++(0, 1.2)
    -- ++(4.0, 0)
    -- ++(0, 15.0)
    -| node[annot, pos=0.10, above]{advance to chunk $t{+}1$} (chunk.north);

  % ---- ROW 6: task head + reward ----
  \node[head]   (taskhead) at (9, -19.0)
    {Task Head\\(cls.\ / detection)};
  \node[reward] (reward)   at (9, -21.5)
    {$R = \mathrm{TaskMetric}$\\
     $-\,\lambda\max(0,1{-}B_T)$};

  \draw[larrow] (mem_new)  -- node[annot, right]{$t=T$} (taskhead);
  \draw[larrow] (taskhead) -- (reward);

  \draw[darrow, bend right=38]
    (reward.west) to
    node[annot, left]{PPO update (GAE)} (actor.west);

  % ============================================================
  % Background phase shading
  % ============================================================
  \begin{scope}[on background layer]
    \node[draw, dashed, rounded corners, fill=gray!3, draw=gray!30,
          fit=(fisher)(actor)(wannot), inner sep=8pt] (p0box) {};
    \node[annot, above right=2pt and 2pt of p0box.north west]
      {\textbf{Phase 1}};

    \node[draw, dashed, rounded corners, fill=red!2, draw=red!18,
          fit=(actor)(wannot)(soft)(hard)(dilution)(active)%
              (mem_new)(taskhead)(reward),
          inner sep=8pt] (p1box) {};
    \node[annot, above right=2pt and 2pt of p1box.north west]
      {\textbf{Phase 2}};

    \node[draw, dashed, rounded corners,
          fill=orange!4, draw=orange!30,
          fit=(hard)(active), inner sep=5pt] {};
    \node[annot, above right=1pt and 1pt of hard.north west]
      {inference only};
  \end{scope}

  \end{tikzpicture}%
  }% end resizebox
  \caption{%
    Architecture overview. An input image is tokenised into $N$ patches and
    processed in $T$ sequential chunks. At each step $t$, the PPO Actor-Critic
    observes the causal state $s_t$ encoding local features, projections of
    the recurrent memory $(S_{t-1}, Z_{t-1})$, and the remaining budget $B_t$
    and emits per-token weights $w_t$. These soft-scale token contributions
    during training; at inference, weights below threshold $\tau$ are
    hard-pruned for exact FLOP reduction. State Dilution replenishes the memory
    with the historical mean $S_{t-1}/(t{-}1)$ proportional to suppression
    $(1-\bar{w}_t)$. After $T$ chunks, the final memory $(S_T, Z_T)$ feeds a
    task head and a terminal reward drives PPO via GAE\@.
    Phase 1 bootstraps the Actor from Fisher Information saliency targets before
    Phase 2 end-to-end fine-tuning.%
  }
  \label{fig:architecture}
\end{figure*}

\subsection{Chunkwise Formulation of Linear Attention}
Standard linear attention leverages the associative property to reduce complexity to $\mathcal{O}(N)$. The output for the $i$-th query is typically computed using a global normalization term:
\begin{equation}
    O_i = \frac{\phi(Q_i) \sum_{j=1}^N \phi(K_j)^\top V_j}{\phi(Q_i) \sum_{j=1}^N \phi(K_j)^\top}
\end{equation}

To enable sequential importance assignment, we reformulate this into a causal, chunkwise process. The input sequence of $N$ tokens is partitioned into $T$ non-overlapping sequential chunks of size $C$. At each step $t \in \{1, \dots, T\}$, the backbone processes the current chunk into local Queries ($Q_t$), Keys ($K_t$), and Values ($V_t$).

Rather than computing a full sequence matrix, we maintain two recurrent memory representations to capture global context. The numerator state $S_t \in \mathbb{R}^{D \times D}$ accumulates the cross-covariance of keys and values, while the denominator state $Z_t \in \mathbb{R}^{D \times 1}$ tracks the normalization factor:
\begin{equation}
    S_t = S_{t-1} + \phi(K_t)^\top V_t
\end{equation}
\begin{equation}
    Z_t = Z_{t-1} + \sum_{i \in \text{chunk } t} \phi(K_{i, t})^\top
\end{equation}

The attention output for the current chunk $t$ is then computed as:
\begin{equation}
    O_t = \frac{\phi(Q_t) S_t}{\phi(Q_t) Z_t}
\end{equation}

\subsection{Markov Decision Process Formulation}

By reformulating linear attention into a recurrent, chunkwise execution, we isolate the processing of each image segment into a discrete-time Markov Decision Process (MDP). This structure allows a routing agent to sequentially determine the importance of tokens based on local information, accumulated global memory, and remaining computational resources.

\subsubsection{State Space ($\mathcal{S}$)}
At each chunk step $t$, the agent observes a causal state $s_t$ defined as:
\begin{equation}
s_t = \left[ K_t \parallel \mathrm{Proj}(S_{t-1}) \parallel \mathrm{Proj}(Z_{t-1}) \parallel B_t \right]
\end{equation}

The components are:
\begin{itemize}
    \item \textbf{Local Features ($K_t \in \mathbb{R}^{C \times D}$):} Visual features of the current chunk.
    
    \item \textbf{Memory Projection ($\mathrm{Proj}(S_{t-1}) \in \mathbb{R}^{D}$):} A learned projection of the accumulated memory matrix $S_{t-1} \in \mathbb{R}^{D \times D}$.
    
    \item \textbf{Normalization Projection ($\mathrm{Proj}(Z_{t-1}) \in \mathbb{R}^{D}$):} A projection of the normalization accumulator $Z_{t-1}$, encoding the effective mass of stored tokens.
    
    \item \textbf{Dynamic Budget ($B_t \in [0,1]$):} Remaining computational budget.
\end{itemize}

\subsubsection{Action Space ($\mathcal{A}$)}
For each chunk of size $C$, the Actor $\pi_\theta$ outputs:
\begin{equation}
a_t = \{ w_{1,t}, w_{2,t}, \dots, w_{C,t} \}, \quad w_{i,t} \in (0,1)
\end{equation}
\subsubsection{Transition Dynamics}
The transition to $s_{t+1}$ is governed by deterministic updates of the memory and normalization accumulators.

The memory update is:
\begin{equation}
S_t = S_{t-1} + \left( (w_t \odot \phi(K_t))^\top V_t \right)
\end{equation}

The normalization term evolves as:
\begin{equation}
Z_t = Z_{t-1} + \sum_{i=1}^{C} w_{i,t} \, \phi(K_{i,t})^\top ,
\end{equation}
where $\phi(\cdot)$ is the feature map associated with linear attention, and $\odot$ denotes element-wise scaling of token features by their corresponding weights.

The budget follows:
\begin{equation}
B_{t+1} = B_t - \frac{1}{N} \sum_{i=1}^{C} w_{i,t}
\end{equation}

\subsubsection{Reward Function ($\mathcal{R}$)}
A terminal reward $R$ is provided at $t = T$:
\begin{equation}
R = \mathrm{TaskMetric} - \lambda_{\text{penalty}} \cdot \max(0, 1 - B_T)
\end{equation}

where:
\begin{itemize}
    \item \textbf{TaskMetric:} Task-specific performance (e.g., mAP or accuracy).
    \item \textbf{Budget Penalty:} Penalizes exceeding the allocated token budget.
\end{itemize}

This formulation encourages the agent to preserve informative tokens while respecting computational constraints.

\subsection{Squashed Gaussian Actor-Critic Network}
To provide the flexibility to down-weight or entirely omit uninformative chunks, we model the routing policy $\pi_\theta(a_t \mid s_t)$ as a Squashed Gaussian distribution. This independent weighting avoids the mathematical instability of discrete gates and the competitive constraints of a global Softmax operation, allowing the network to allocate exact importance independent of the surrounding context. 

The Actor network computes the parameters for an unbounded Normal distribution. To ensure the standard deviation remains positive and numerically stable, it is passed through a Softplus activation with a small $\epsilon$ floor:
$$ \mu_\theta = \text{MLP}_\mu(s_t), \quad \sigma_\theta = \text{Softplus}(\text{MLP}_\sigma(s_t)) + \epsilon $$

During stochastic training, actions are sampled using the reparameterization trick, allowing gradients to flow back through the stochastic node. The unbounded sample $x$ is then squashed via a hyperbolic tangent and shifted to the valid continuous weighting space of $(0, 1)$:
$$ w_t = \frac{\tanh(x) + 1}{2}, \quad x \sim \mathcal{N}(\mu_\theta, \sigma_\theta^2) $$

By decoupling the weights from a global sum, the Actor can assign near-zero weights to every token in a redundant chunk, effectively "muting" its contribution to the global memory $S_t$. However, applying the non-linear Tanh transformation changes the probability density of the sampled action. To maintain mathematical integrity for Proximal Policy Optimization (PPO), the log-probability of the action must be corrected using the Jacobian trace of the Tanh transformation:
$$ \log \pi(w_t) = \log p(x) - \sum \log(1 - \tanh^2(x) + \epsilon) $$

Concurrently, a Value network (Critic) $V_\phi(s_t)$ evaluates the causal state to estimate the expected episodic return, which serves as a baseline to stabilize the subsequent RL updates.

\subsection{Pretraining: Scale-Aligned Information Distillation}
Initializing an RL agent from scratch in a massive parameter space often leads to representation collapse, as the agent initially acts randomly and destroys the transformer's memory state. To circumvent this, Phase 1 initializes the Actor using empirical Fisher Information ($\mathcal{I}_i$) as a saliency oracle. 

After warming up the backbone on a dense classification task, we compute the squared gradients of the cross-entropy loss with respect to the token features to approximate their Fisher Information. These scores reflect the true informational utility of each token. To prevent scale shift during the transfer to the RL phase, we apply spatial smoothing and normalize these scores to a strict $\tilde{\mathcal{I}}_i \in [0, 1]$ range.

The Actor and Critic are then trained via supervised learning. The Actor minimizes the Mean Squared Error between its deterministic mean prediction $w_{\mu}$ and the Fisher utility, while the Critic learns to predict the total saliency of the chunk:
$$ \mathcal{L}_{\text{Actor}} = \mathbb{E} \left[ \left( w_{\mu}(s_t) - \tilde{\mathcal{I}}_t \right)^2 \right] $$
$$ \mathcal{L}_{\text{Critic}} = \mathbb{E} \left[ \left( V_\phi(s_t) - \sum_{i \in t} \tilde{\mathcal{I}}_i \right)^2 \right] $$

\subsection{Finetuning: Task-Specific Reward Optimization}
In Phase 2, the model is transitioned to the downstream task and optimized via PPO. The environment dynamically enforces a decreasing budget $B_t$, forcing the Actor to plan its token allocation temporally. 
Aggressive down-weighting can cause the recurrent memory to stagnate, leading to 
catastrophic forgetting of accumulated visual context. To mitigate this, the memory is augmented at each step by a fraction of its own historical average, scaled by the degree of suppression $(1 - \bar{w}_t)$:
\begin{equation}
    S_t = S_{t-1} + \left(w_t \odot \phi(K_t)\right)^\top V_t 
    + \gamma \cdot (1 - \bar{w}_t) \cdot \frac{S_{t-1}}{t-1}
\end{equation}
where the dilution term is defined as zero for $t=1$. When the agent fully suppresses 
a chunk $({\bar{w}_t \to 0})$, the memory is topped up with the mean KV contribution 
accumulated so far, preventing stagnation. When all tokens are retained 
$(\bar{w}_t \to 1)$, the term vanishes and the update reduces to standard linear 
attention.

Using Generalized Advantage Estimation ($\hat{A}_t$) derived from the Critic's value predictions, the Actor is updated via the clipped PPO surrogate objective. To guarantee sufficient exploration of the continuous weight space and prevent premature convergence, we inject an entropy bonus:
\begin{equation}
\begin{split}
\mathcal{L}_{\mathrm{PPO}} = &-\mathbb{E}\!\left[\min\!\left(r_t(\theta)\hat{A}_t,\;
    \mathrm{clip}(r_t(\theta),1\pm\epsilon)\hat{A}_t\right)\right] \\
&- \lambda_{\mathrm{ent}}\, H(\pi_\theta)
\end{split}
\end{equation}
where

\begin{equation}
H(\pi_\theta) = \frac{1}{2}\log(2\pi e\,\sigma_\theta^2) 
+ \mathbb{E}\!\left[\sum_i \log(1 - \tanh^2(x_i) + \epsilon)\right]
\end{equation}
is the entropy of the squashed Gaussian policy.

During inference, stochastic sampling is disabled. The Actor deterministically outputs a continuous weight $w_{\text{infer}}$. A hard threshold $\tau$ enables physical token purging, stripping uninformative tokens from the GPU memory buffer to yield exact FLOP reductions. Crucially, surviving tokens retain their continuous scaling factors $w_{\text{infer}}$ in the linear attention update, bypassing the quantization shock typically associated with discrete gating mechanisms.